### 2020-09-21 LaTeX/sections/appendix.tex
Appendix added

% Extended discussion on cleaning up data structures in Ethereum

\section{Ethereum's Gas Model}

Gas is the Ethereum's pseudo-currency and a key variable for the execution of smart contracts. DApps written in high-level programming languages are compiled and translated into a compact representation (called ‘bytecode’) to be further executed on the Ethereum virtual machine or EVM. Each of these opcodes has a fixed amount of gas assigned and is a measure of computational effort. Separating ETH from gas prevents transactions cost from being too expensive as they are less impacted by ETH price fluctuations. Gas unit does not a have a monetary value and as mentioned, it only  measures the computational work undertaken by miners. To pay miners, Ethereum introduces a \texttt{gas price} --- small denomination of ether called \textbf{Gwei} attached to each gas unit. Essentially, gas price indicates how much users are willing to pay per unit of gas, clearly the higher gas price leads to the faster execution of a transaction. Miners also need to know the total amount of computational work a user is requesting, called the \texttt{gas limit}. \texttt{Gas limit} is a parameter that limits the amount of gas users would spend in a transaction. It protects users from spending unlimited ETH on their transactions and must be set carefully. If the \texttt{gas limit} is too low, the transaction will exceed the limit, all operations will be reverted while the user must pay for the computational work performed by the miner. By default the gas limit for an Ethereum transaction is 21,000 gas. An Ethereum block also has a \texttt{block gas limit} field which is set by the Ethereum miners and indicates the maximum amount of gas all the transactions in that block are allowed to consume, the Ethereum \texttt{block gas limit} is currently 11,741,495. \footnote{July 2020: https://ethstats.net/}

\subsection*{Gas Refunds.}There are two particular EVM operations with negative gas --- certain amount of gas is refunded to the sender at the end of the transaction. These operation include:

\paragraph{\texttt{SELFDESTRUCT}.}This operation destroys and deletes the originating contract and refunds its balance (if any) to a designated receiver address. Note that the Ethereum storage is implemented in the form of a has map and EVM is not statistically aware of which storage slots are held by the contract. Because of this the \texttt{SELFDESTRUCT} operation does not remove the initial byte code of the contract from the chain, but it frees up the state storage and has a refund of 24,000 gas.

@@

Note that in order to urge miners to process smart contracts with refunds, the accumulated gas refund can never exceed half the gas used up during computation~\cite{wood2014ethereum}. So at the end of a successful transaction, the amount of gas in the refund counter (capped at half the net gas used) is returned to the caller. At the time of this writing, Ethereum transaction receipts only account for the \texttt{gasUsed}, which is the total amount of gas units spent during a transaction, and we are not able to obtain the value of the EVM's refund counter from inside the EVM~\cite{signer2018gas}. So in order to account for refunds, we decide to calculate them manually; first we figure out exactly how much storage is being cleared or how many smart contracts are being destroyed, then we multiply these numbers by 24,000 and 15,000 respectively.

@@

To maintain the collateral balance of each market participant, we use two Solidity type one-to-one mappings that map Ethereum addresses to 256 bits unsigned integers; \texttt{TotalBalance} and \texttt{UnavailableBalance}. Once market closes and orders are matched, the \texttt{UnavailableBalance} needs to be cleared. However, since it is not possible to delete the entire mapping without knowing the keys~\footnote{https://solidity.readthedocs.io/en/v0.5.12/security-considerations.html}, clearing the \texttt{UnavailableBalance} mapping remains a challenging issue to solve. Here we provide a landscape of solutions for that.
%only individual keys and what they map to can be deleted.

\begin{itemize}

%---------+++++++++++++++++++-----------------
\item \textbf{Creating a New Mapping Every Time the Market Opens.} Instead of clearing the \texttt{UnavailableBalance} of traders at the end of a matching process, we could create a new mapping every time the market opens. Note that in this scheme traders can claim their funds back (using the \texttt{ClaimEther()} and/or \texttt{ClaimToken} methods) only when the market is closed (\ie the market is in the \texttt{Closed} state).

%---------+++++++++++++++++++-----------------
\item \textbf{Creating Custom Keys for the Mapping.} We can create custom keys for the mapping by defining a counter as a global variable inside the call market smart contract and increment it only once at the end of the matching process. So instead of clearing the mapping, we only use another portion of it every time the market opens.

%---------+++++++++++++++++++-----------------
\item \textbf{Storing the Mapping in a Data-Contract.} Another design proposal is to create a new smart contract every time the market opens, this data contract only stores the \texttt{UnavailableBalance} mapping and will be destroyed (using the \texttt{SELFDESTRUCT} operation) at the end of the matching process.

%---------+++++++++++++++++++-----------------
\item \textbf{Storing the Mapping keys in an additional Array.} Another common pattern is to create an additional array on top of the mapping and iterate over that. This array (\eg address[]) stores the traders' addresses and enables us to iterate over the mapping and delete individual keys and what they map to at the end of the matching process. Note that this is a gas-costly design pattern as we would need to maintain a secondary data structure.

%---------+++++++++++++++++++-----------------
\end{itemize}

>>>>

\section{Extended Discussion on Cleaning up Data Structures on Ethereum}

As discussed in Section~\ref{sec:gasrefund}, whether to clean up storage variables or not remains a vital challenge for developers when developing DApps on the Ethereum blockchain--- this is even more crucial for implementation of \cm, as we needed  to choose the best middle ground design model which gives us a satisfactory performance and cost results. For this reason, we believe it it is helpful (both for researchers and developers) to extend our discussion on this topic. First, we discuss the Ethereum's gas model--- (i) what are different gas parameters one needs to take into account when sending a transaction on the Ethereum blockchain, (ii) what is the gas refund, and (iii) and how it is calculated.  Next, we discuss one major challenge, \textit{clearing mappings}, that we faced when implementing \cm on the Ethereum.

\subsection{Ethereum's Gas Model}

Gas is the Ethereum's pseudo-currency and a key variable for the execution of smart contracts. DApps written in high-level programming languages are compiled and translated into a compact representation (called ‘bytecode’) to be further executed on the Ethereum virtual machine or EVM. Each of these opcodes has a fixed amount of gas assigned and is a measure of computational effort. Separating ETH from gas prevents transaction costs from being too expensive as they are less impacted by ETH price fluctuations. Gas unit does not a have a monetary value and as mentioned, it only  measures the computational work undertaken by miners. To pay miners, Ethereum introduces a \texttt{gas price} --- small denomination of ether called \textbf{Gwei} attached to each gas unit. Essentially, gas price indicates how much users are willing to pay per unit of gas--- clearly a higher gas price leads to the faster execution of a transaction. Miners also need to know the total amount of computational work a user is requesting, called the \texttt{gas limit}. \texttt{Gas limit} is a parameter that limits the amount of gas users would spend in a transaction, it protects users from spending unlimited ETH on their transactions and must be set carefully. If the \texttt{gas limit} is too low, the transaction will exceed the limit and all operations will be reverted while the user must pay for the computational work performed by the miner. By default the gas limit for an Ethereum transaction is 21,000 gas. An Ethereum block also has a \texttt{block gas limit} field which is set by the Ethereum miners and indicates the maximum amount of gas all the transactions in that block are allowed to consume, the Ethereum \texttt{block gas limit} is currently 11,741,495. \footnote{July 2020: https://ethstats.net/}

\subsubsection*{Gas Refunds.}There are two particular EVM operations with negative gas--- certain amount of gas is refunded to the sender at the end of the transaction. These operations include:

\paragraph{\texttt{SELFDESTRUCT}.}This operation destroys the originating contract and refunds its balance (if any) to a designated receiver address. Note that the Ethereum storage is implemented in the form of a has map and EVM is not statistically aware of which storage slots are held by the contract. So the \texttt{SELFDESTRUCT} operation does not remove the initial byte code of the contract from the chain, but it frees up the state storage and has a refund of 24,000 gas.

@@

Note that in order to urge miners to process smart contracts with refunds, the accumulated gas refund can never exceed half the gas used up during computation~\cite{wood2014ethereum}. So at the end of a successful transaction, the amount of gas in the refund counter (capped at half the net gas used) is returned to the caller. At the time of this writing, Ethereum transaction receipts only account for the \texttt{gasUsed}, which is the total amount of gas units spent during a transaction, and users are not able to obtain the value of the EVM's refund counter from inside the EVM~\cite{signer2018gas}. So in order to account for refunds, we can only calculate them manually; first we need to figure out exactly how much storage is being cleared or how many smart contracts are being destroyed, then we multiply these numbers by 24,000 and 15,000 respectively.

@@

To maintain the ETH and ERC20 tokens balances of each trader in \cm, we use one-to-one mappings, \texttt{totalBalance} and \texttt{unavailableBalance}, that map Ethereum addresses to unsigned integers. Note that we need to clear up the entire  \texttt{unavailableBalance} mapping during the orders settlement and clearing phase. However in Solidity, it is not possible to delete the entire mapping without knowing the keys~\footnote{In a Solidity mapping, only individual keys and what they map to can be deleted: https://solidity.readthedocs.io/en/v0.5.12/security-considerations.html}. Here we propose and discuss the landscape of various solutions together with how each method affects the total amounts of gas costs for clearing orders in \cm.

\begin{enumerate}

%---------+++++++++++++++++++-----------------
\item \textbf{Storing mapping keys in additional array.} This is a common design pattern used by DApp developers--- to delete the mapping, one needs to create an additional array on top of it and iterate over that array. For example in \cm, we can create \texttt{address[] helperArray} that stores traders' addresses. Knowing traders addresses (\ie mapping keys), we can now iterate over the mapping and delete these individual keys and what they map to at the end of the matching/clearing process. Despite being a common design pattern, this solution 
significantly increases the total cost of matching the orders as Solidity loops are costly operations.

%---------+++++++++++++++++++-----------------
\item \textbf{Storing the mapping in a separate smart contract.} Another design proposal is to create a new smart contract every time the market opens. This contract \textit{only} stores the \texttt{unavailableBalance} mappings and will be destroyed (using the \texttt{SELFDESTRUCT} operation) at the end of the matching process. This solution assigns overhead costs to the \texttt{matchOrders()} function as it involves an external call to another smart contract (\ie destroy the contract containing the mappings).

%---------+++++++++++++++++++-----------------
\item \textbf{Generating custom mapping keys.} We can also define a global counter variable and use it as a custom key for the mappings. This counter is incremented \textit{only once} at the end of the matching process. So we only use a new portion of the mapping (and leave the used portion on the Ethereum blockchain) every time the market opens. This solution removes the burden of the gas cost from the \texttt{matchOrders()} function (as discussed in Section~\ref{sec:gasrefund}, leaving the data on chain is cheaper then cleaning it up), however DApp developers are advised to be good citizens of Ethereum and cleanup on-chain data they no longer need.

%---------+++++++++++++++++++-----------------
\item \textbf{Creating a new mapping every time the market opens.} Similar to the previous proposal, we can skip deleting mappings and instead create a new mapping every time the market opens. Again, although this method does not increase the cost of matching the orders in \cm, it leaves unused data on the Ethereum blockchain.

%---------+++++++++++++++++++-----------------
\end{enumerate}
